---
title: "Lab 2"
author: "Sierra Jo Sallee"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 5
    number_sections: no
    toc: yes
    toc_float: yes
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```
# Task 1
Get working directory.
<br>
```{r}
getwd()
```
# Task 2
Read data into R.
<br>
```{r}
dataTable = read.csv("/Users/sierrasallee/Desktop/Stats/Labs/DATAxls/EPAGAS.csv")

head(dataTable)
```
# Task 3
Create the variable mpg
<br>
```{r}
mpg = dataTable$MPG

zi = function(x){ # x is a vector
  n <- length(x)
  x_mean <- mean(x)
  (x - x_mean) / sd(x)
}
```
<br>
Z-transformation of mpg
<br>
```{r}
z = zi(mpg)
z
```
<br>
Mean value of z
<br>
```{r}
z_mean = round(mean(zi(mpg)), digits=2)
z_mean
```
<br>
Standard deviation of z squared
<br>
```{r}
sz2 = sd(zi(mpg))^2
sz2
```
<br>
Possible outliers
<br>
```{r}
possible_outliers = mpg[abs(z)>=2&abs(z)<=3]
possible_outliers[!is.na(possible_outliers)]
```
<br>
Outliers
<br>
```{r}
outliers = mpg[abs(z)>3]
outliers[!is.na(outliers)]
```
<br>
Dotplot of possible outliers, outliers, and main line values.
<br>
```{r}
z_value = abs(round(z, digits=0))

dotplot_data = data.frame(dataTable, z_value)
ggplot(dotplot_data, aes(x=" ", y=MPG, fill=factor(z_value))) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  ggtitle("Sierra Jo Sallee") 
```

# Task 4
Make a boxplot of the mpg variable
<br>
```{r}
dataTable = read.csv("/Users/sierrasallee/Desktop/Stats/Labs/DATAxls/EPAGAS.csv")

ggplot(dataTable, aes(x=" ", y=MPG)) + 
  geom_boxplot(notch=TRUE, fill="black") +
  coord_flip() +
  ggtitle("Sierra Jo Sallee")
```
<br>
Using Chebyshev’s theorem predict the proportion of data within 2 standard deviations of the mean of the mpg data.
<br>
```{r}
length(mpg) * .75
```
<br>
Use R to calculate the exact proportion within 2 standard deviation of the mean.
<br>
```{r}
c_high = mean(mpg) + (2 * sd(mpg))
c_low = mean(mpg) - (2 * sd(mpg))

length(with(dataTable, dataTable[MPG <= c_high & MPG >= c_low,]))

```
<br>
Does Chebyshev agree with the data? <br>
Yes. Chebyshev's rule dictates that AT LEAST 75% of the data will be within 2 standard deviations of the mean. Our data has more than 75%, but it still follows Chebyshev's prediction. <br>
<br>
Now use the empirical rule, what proportion (according to the rule) of the data should be within 2 standard deviations of the mean?
<br>
```{r}
length(mpg) * .95
```
<br>
How well does it correspond?
Ther is only 1% of difference in the data from the prediction. It corresponds well.
<br>
Is the Empirical rule valid in this case? Why?
Yes. The Empirical rule states that APPROXIMATELY 95% of the data will be within 2 standard deviations of the mean. 96% is approximately 95%.